Multi-Core Model Checking and Maximum Satisfiability Applied to Hardware-Software Partitioning

### 2016-09-25 Docs/IJES/journal.tex
final version - rev1

%\LRH{A. Trindade et~al.}

@@

\PUBYEAR{xxxx}

@@

\PUBYEAR{2012}

@@

\authorA{Author A}
\affA{University A\\
E-mail: \{nonono\}@nonono.com}

\authorB{Author B}
\affB{University B\\
E-mail: \{nonono\}@nonono.com}

%\authorA{Alessandro Trindade, Hussama Ismail}
%\affA{Federal University of Amazonas \\
%E-mail: \{alessandro.b.trindade, hussamaismail\}@gmail.com}

%\authorB{Edilson Galv\~ao, Renato Degelo}
%\affB{Federal University of Amazonas \\
%E-mail: \{rdegelo, esj.galvao\}@gmail.com}

%\authorC{Helder Silva, and Lucas Cordeiro}
%\affC{Federal University of Amazonas \\
%E-mail: prhsilva2012@gmail.com, lucascordeiro@ufam.edu.br}

\begin{abstract}
\textcolor{blue}{Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) is well known by its capability to verify software. However, its use as optimization tool, to solve hardware and software (HW-SW) partitioning problem, is something new. In particular, its integration with the Maximum Satisfiability solver $\nu$Z tool, which provides a portfolio of approaches for solving linear optimization problems over SMT formulas, is unprecedented. We present new alternative approaches to solve the HW-SW partitioning problem. First, we use SMT-based BMC in conjunction with a multi-core support using Open Multi-Processing to create four variants to solve the partitioning problem.
The multi-core SMT-based BMC approaches allow initializing many verification instances based on the number of available processing cores, where each instance checks a different optimum value until the optimization problem is satisfied. Additionally, we integrate the $\nu$Z into the BMC, making it as a specialized solution for optimization in a single-core environment. We implement all five approaches on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC) and compare them to a state-of-the-art optimization tool (MATLAB). Experimental results show that there is no single optimization tool to solve all HW-SW partitioning benchmarks, but based on medium-size benchmarks, ESBMC-$\nu$Z had better performance.}

@@

\REF{\textcolor{red}{to this paper should be made as follows: No no no. `Multi-Core Model Checking and Maximum Satisfiability Applied to Hardware-Software Partitioning', {\it International Journal of Embedded Systems}, Vol. x, No. x, pp.xxx\textendash xxx.}}

%Trindade, A., Ismail, H., Galv\~ao, E., Degelo, R., Silva, H., and Cordeiro, L.%
\begin{bio}
\textcolor{red}{
no no no.\vs{9}}

\textcolor{red}{\noindent No no no.\vs{8}}

\textcolor{red}{\noindent No no no.}
\end{bio}

@@

%-----------------related work came to here and was rewriten
\textcolor{blue}{Since the first decade of 2000s, two main paths have been tracked to solve the HW-SW partitioning problem: to find the exact solution of a optimization problem, as presented by~\cite{Mann2007}; and to use heuristics to speed up performance time, as presented by~\cite{Arato2003} and~\cite{Arato2005}. It is worth mentioning that using heuristics the final solution is not necessarily an optimal global solution.}

\textcolor{blue}{There is also the effort to create a hardware-software (HW/SW) co-simulation environment in order to help designers obtain an appropriate HW/SW partitioning that satisfy specified tradeoffs, as described by~\cite{Jan2005} and by~\cite{Hau2009}.}

\textcolor{blue}{In terms of SMT-based verification, most related studies are restricted to present the model, its modification to specific programming languages ({\it e.g.}, C/C++ and Java), and the application to multi-thread algorithms or to embedded systems to check for program correctness. In~\cite{Ramalho2013} it presents a bounded model checker for C++ programs, which is an evolution of dealing with C programs, and~\cite{Cordeiro2012} use ESBMC for embedded ANSI-C software. In~\cite{Trindade2015}, and~\cite{Trindade2016} it was proven that it is possible to use ESBMC to solve HW-SW partitioning in a single- and multi-core way, but the former has performance issues that were improved by the latter, which used only a sequential search to perform multi-core model checking. There are related studies focused on decreasing the verification time of model checkers by applying Swarm Verification, as mentioned in~\cite{Holzmann2011}, and modifications of internal search engines to add support for parallelism, as presented by~\cite{Holzmann2012}, but there is still the need for initiatives related to parallel SMT solving, as described by~\cite{Wintersteiger2009}.}

\textcolor{blue}{Recently, the SMT solver Z3 has been extended to pose and solve optimization problems modulo theories, as presented by~\cite{Bjorner2015}. In particular, $\nu$Z tool offers substantial performance improvement in optimization problems, as described by~\cite{Bjorner2014}. Additionally, as an application example, \cite{Pavlinovic2015} propose an approach which considers all possible compiler error sources for statically typed functional programming languages and reports the most useful one subject to some usefulness criterion. The authors formulate this alternative single-core approach as an optimization problem related to SMT and use $\nu$Z to compute an optimal error source in a given ill-typed program.}

@@

\textcolor{blue}{Here, we apply SMT-based verification methods to the HW-SW partition problem in three different ways using a multi-core ESBMC approach with OpenMP: ESBMC-SS using a sequential-search (SS), ESBMC-PS using a parallel-search (PS), and ESBMC-PB using a binary-search (BS). Our experimental results are compared to ILP (integer linear programming) and GA (generic algorithms) in a multi-core version, and also to $\nu$Z, which supports only a single-core approach, as described by~\cite{Bjorner2015}. The ILP and GA algorithms are implemented with the optimization toolbox of Matlab, as described in~\cite{TheMathWorks2013}, while $\nu$Z is a built-in tool to the SMT solver Z3. All multi-core ESBMC approaches, together with $\nu$Z, are implemented with the ESBMC tool. To the best of our knowledge, this is the first work to use a multi-core SMT-based verification and a MaxSMT solver to check for HW-SW partitioning problems in embedded systems.}

%-----------------
%Starting at the $90$'s decade, intensive research was performed in HW-SW partitioning, and several approaches proposed, as shown in~\cite{Arato2003} and~\cite{Mann2007}. In~\cite{Trindade2015} and~\cite{Trindade2016} was shown that it is possible to use Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT), implemented in a tool called Efficient SMT-Based Context-Bounded Model Checker (ESBMC) as defined in~\cite{Cordeiro2012}, in order to perform HW-SW partitioning in embedded systems. In~\cite{Trindade2016}, the original ESBMC algorithm was improved in terms of performance, so a multi-core sequential approach was introduced, in order to speed up the performance and turn it similar, in some cases, to conventional optimization tools, which are used to perform HW-SW partitioning.

%The present work thus extends those previous studies since there is a substantial improvement in terms of the SMT-based verification methods. In particular, we extend the work of~\cite{Trindade2015} and~\cite{Trindade2016} to exploit a multi-core parallel- and binary-search approaches, as well as, the integration of the Maximum SMT (MaxSMT) solver $\nu$Z into ESBMC, which is a state-of-art optimization \textcolor{blue}{single-core} tool based on SMT as defined in~\cite{Bjorner2015}.

%We exploit the availability of multi-core processors and the tool OpenMP as defined in~\cite{Tang2009}; thus, we apply SMT-based verification methods to the HW-SW partition problem in three different ways using a multi-core ESBMC approach with OpenMP: ESBMC-SS using a sequential-search (SS), ESBMC-PS using a parallel-search (PS), and ESBMC-PB using a binary-search (BS). Experimental results are compared to ILP (integer linear programming), GA (generic algorithms) in a multi-core version, and also to $\nu$Z, which supports only a single-core approach, according to~\cite{Bjorner2015}. The ILP and GA algorithms are implemented with the optimization toolbox of Matlab, according to~\cite{TheMathWorks2013}, while $\nu$Z is a built-in tool to the SMT solver Z3. All multi-core ESBMC approaches, together with $\nu$Z, are implemented with the ESBMC tool.

%The main contribution of the present study is to describe and evaluate a comprehensive SMT-based BMC approach in a multi-core architecture applied to solve HW-SW optimization problems. Additionally, we integrate the MaxSMT solver $\nu$Z into an off-the-shelf BMC tool, which is typically used for software verification, in order to formulate and solve optimization problems within the logical context of constraints. Experimental results show that multi-core model-checking techniques can be effective, in particular cases, to find the optimal solution of the HW-SW partitioning problem using an SMT-based BMC approach. Although there is no single tool for efficiently solving all HW-SW partitioning problems, we show that the MaxSMT solver $\nu$Z is faster than other state-of-the-art optimization tools for small- and medium-size optimization problems.  To the best of our knowledge, this is the first work to use a multi-core SMT-based verification and a MaxSMT solver to check for HW-SW partitioning problems in embedded systems.

%\subsection{Contributions}

%The main contribution of the present study is to describe and evaluate a comprehensive SMT-based BMC approach in a multi-core architecture applied to solve HW-SW optimization problems. Additionally, we integrate the MaxSMT solver $\nu$Z into an off-the-shelf BMC tool, which is typically used for software verification, in order to formulate and solve optimization problems within the logical context of constraints. Experimental results show that multi-core model-checking techniques can be effective, in particular cases, to find the optimal solution of the HW-SW partitioning problem using an SMT-based BMC approach. Although there is no single tool for efficiently solving all HW-SW partitioning problems, we show that the MaxSMT solver $\nu$Z is faster than other state-of-the-art optimization tools for small- and medium-size optimization problems.  To the best of our knowledge, this is the first work to use a multi-core SMT-based verification and a MaxSMT solver to check for HW-SW partitioning problems in embedded systems.

@@

Our experiments are based on a set of publicly available benchmarks. \textcolor{blue}{All tools, benchmarks, and results of our evaluation are available on a supplementary web page \texttt{http://esbmc.org} and \texttt{http://esbmc.org/benchmarks}}.

@@

Optimization is the act of obtaining the best result ({\it i.e.}, the optimal solution) under given circumstances as defined in~\cite{Rao2009}. There is no single method available for efficiently solving all optimization problems, as described by~\cite{Rao2009}. The most well-known technique is linear programming, which is a method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. \textcolor{blue}{Eq.~\eqref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors from the objective function, while $Aeq$ and $beq$ are matrices that describe the linear equality constraints}

@@

\vspace{2 mm}
\textcolor{blue}{Here, we refer to the relevant work of ~\cite{Mann2007} in optimization, who modified a branch-and-bound algorithm to speed up the execution time. ~\cite{Sapienza2013} used multiple criteria decision analysis to solve the HW-SW partitioning problem.}

In some cases, the time to find a solution using ILP is impractical. Even with the use of powerful computers, a problem can take hours before an optimal solution is reached. If the optimization problem is complex, some heuristics can be used to solve the same problem faster, as described by~\cite{Rao2009}, {\it e.g.}, those used in the GA. The only drawback is that the found solution may not be the global minimum or maximum. \textcolor{blue}{As relevant studies, ~\cite{Jiang2012} used a GA mixed with simulated annealing and~\cite{Huong2012} designed a modified Pareto optimization using GA. In the other hand, there is some research done to reduce the complexity of computation in order to improve the performance of some techniques, as described by~\cite{Boucheneb2006}.}

\textcolor{blue}{Alternatively, tools such as ESBMC and $\nu$Z can be used to solve optimization problems so that the global minimum or maximum solution is found, as described by~\cite{Trindade2015} and~\cite{Bjorner2015}}. The following sections describe the main features of ESBMC and $\nu$Z tools.

@@

Among recent model checking techniques, there is one that combines model checking with satisﬁability solving. This technique, known as bounded model checking (BMC), does a very fast exploration of the state space, and for specific types of problems, it offers large performance improvements over previous approaches, as \textcolor{blue}{presented by}~\cite{Biere2009}. In particular, BMC based on Boolean Satisfiability (SAT) has been introduced as a complementary technique to binary decision diagrams for alleviating the state-space explosion problem, as described by~\cite{Clarke2001}.

The basic idea of BMC is to check the negation of a given property at a given depth: given a transition system $M$, a property $\phi$, and a bound $k$, BMC unrolls the system $k$ times and translates it into a verification condition (VC) $\psi$  such that $\psi$ is satisfiable if and only if $\phi$ has a counterexample of depth $k$ or less, as defined in~\cite{Biere2009}. To cope with increasing software complexity, SMT solvers can be used as back-ends for solving the generated VCs, \textcolor{blue}{as presented by~\cite{Ganai2006},~\cite{Armando2009},~\cite{Cordeiro2012}}.

@@

The main factor for ESBMC to use only a single-core relies on its back-end ({\it i.e.}, SMT Solver). Currently, the SMT solvers supported by ESBMC are: Z3, as \textcolor{blue}{presented by}~\cite{DeMoura2008}; Boolector, as \textcolor{blue}{mentioned in}~\cite{Brummayer2009}; MathSAT, as \textcolor{blue}{presented by}~\cite{Barrett2011}; CVC4, as \textcolor{blue}{presented by}~\cite{Bozzano2005}; and Yices, as \textcolor{blue}{presented by}~\cite{Dutertre2014}. Most of them do provide neither multi-threaded support nor a parallel version to solve the generated SMT equations.

@@

In particular, OpenMP uses the \textit{fork-join} model of parallel execution, as \textcolor{blue}{mentioned} in~\cite{OpenMP1998}. The main thread executes the sequential parts of the program; if a parallel region is encountered, then it forks a team of worker threads. After the parallel region finishes ({\it i.e.}, the API waits until all threads terminate), then the main procedure returns to the single-threaded execution mode, as \textcolor{blue}{presented by}~\cite{Wu2014}.

@@

In this study, the SMT solver Z3 is used to check for the satisfiability of formulas generated from the HW-SW partitioning problem, as \textcolor{blue}{presented by}~\cite{Bjorner2014}. In particular, we exploit the use of MaxSMT solver $\nu$Z, which is implemented on top of the SMT solver Z3, in order to solve optimization problems; the $\nu$Z main function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.~$\nu$Z also includes an incremental version of the Maximum Resiliency (MaxRes), as \textcolor{blue}{mentioned} in~\cite{Federica2008}, in order to achieve Maximum Satisfiability (MaxSAT), as defined in~\cite{NarodytskaN} and a Simplex to solve numbers without defined patterns.

In $\nu$Z, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives, as \textcolor{blue}{presented by}~\cite{Bjorner2015}. In summary, $\nu$Z provides three main functions that extend Z3 for solving optimization problems, which are: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

\begin{itemize}
\item{\textbf{maximize(T)}
\textcolor{blue}{this function instructs the solver that a given variable $T$ should be maximized, which includes real, integer, or bit-vector variables.}}
\item{\textbf{minimize(T)}
\textcolor{blue}{this function instructs the solver that a given variable $T$ should be minimized; the accepted types are the same as maximize function.}}
\item{\textbf{Assert-Soft F : weight n}
the function \textit{assert-soft} adds a restriction to $F$, which can also add a weight $n$; the default value is $1$.}
\end{itemize}

@@

\vspace{2 mm}
Fig.~\ref{vZ-Architecture} shows the $\nu$Z architecture. Initially, the SMT formula with objectives is converted to $0-1$ constraints, which leads to a Pseudo-Boolean Optimization (PBO), as \textcolor{blue}{mentioned} in~\cite{Barth1995} and~\cite{Vasco2005}. If there are many objective functions, $\nu$Z invokes OptSAT for arithmetic or MaxSAT for soft constraints. For constraints using real values, $\nu$Z combines linear arithmetic objectives and uses only one instance of OptSMT. When ``soft constrains'' is used in the mode ``lexicographic'', $\nu$Z invokes MaxSAT using multiple calls for its engine.

@@

The informal model can be described by five characteristics. First, there is only one software context, {\it i.e.}, there is just one general-purpose processor, and there is only one hardware context. The system's components must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the execution time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, or energy consumption; \textcolor{blue}{the decision is typically made by the engineer or from the project requirement.} Fourth, based on the assumption that hardware is significantly faster than software, the execution time of the components in hardware is considered to be zero. Finally, if two components are mapped to the same context, then there is no communication overhead between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this study. Hardware components do not need scheduling, because the execution time is assumed to be zero. Since there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, \textcolor{blue}{with a simplified HW-SW model,} where the focus is on bipartitioning, as presented by~\cite{Teich2012}.

@@

\textcolor{blue}{A directed simple graph $ G = (V,E) $, called the task graph of the system, is given. Where the vertices $V = \{V_1,V_2,\dotso,V_n\}$ represent the nodes that are the components of the system that will be partitioned. The edges $E$ represent communication between components. Additionally, each node  $V_i$ has a cost $h(V_i)$ (or $h_i$) of hardware if implemented in hardware and a cost $s(V_i)$ (or $ s_i $) of software if implemented in software. Finally, $c(V_i,V_j)$ represents the communication cost between $V_i$ and $V_j$ if they are implemented in different contexts (hardware or software).}

@@

\vspace{2 mm}
\textcolor{blue}{In particular, different optimization (and decision) problems can be defined for partitioning HW-SW, as described by~\cite{Arato2003}. In this paper, however, the focus is on systems with hard real-time constraints: $ S_0 $ is given (initial cost of software), {\it i.e.}, the objective is to find a $P$ HW-SW partitioning so that $ S_p \leq S_0 $ and $ H_p $ is minimal, which is thus related. Therefore, the objective function is:}

@@

\textcolor{blue}{Consequently, the constraints can be reformulated based on Equations~\eqref{linear-programming-problem} and~\eqref{software-communication-costs} as:}

@@

\vspace{2 mm}
\noindent \textcolor{blue}{where $h$, $s$ and $c$ are the vectors representing the cost functions, $E$ is the transposed incidence matrix of $G$ (indicating which edges cross the boundary between the contexts of hardware and software), $n$ is the number of nodes, and $x$ represents the decision variable (a binary vector indicating the partition: $1$ if the node is realized in hardware and $0$ if the node is realized in software)}. Concerning the complexity of this problem, \cite{Arato2003} demonstrate that it is NP-Hard, as defined in~\cite{Cormem}.

@@

\noindent \textcolor{blue}{As computer hardware architecture moves from single- to multi-core, and more recently to heterogeneous computing as described by~\cite{Mittal2015}, parallel programming environments should be exploited to take advantage of the ability to run several threads on different processing cores}. This section describes the verification algorithm using sequential ESBMC, followed by three multi-core model checking algorithms and the integration of the MaxSMT solver $\nu$Z into ESBMC, in order to speed up the HW-SW partitioning verification. HW-SW partitioning using ILP-based and Genetic Algorithms are also explained.

@@

\textcolor{blue}{Note that, as defined by the formal model, there is an index for each decision variable, indicated by the letter \textit{``i''}, which ranges from $0$ to the number of nodes of the (particular) problem to be solved.}

@@

\textcolor{blue}{In the ESBMC algorithm, there is no need for adding slack variables in Eq.~\eqref{hw-sw-partitioning}, which reduces the number of variables to be solved if compared to ILP and GA.}

>>>>

\LRH{A. Trindade et~al.}

@@

\PUBYEAR{2016}

@@

\PUBYEAR{2016}

@@

\authorA{Alessandro Bezerra Trindade*, Renato de Faria Degelo, Edilson Galv\~ao dos Santos Junior, Hussama Ibrahim Ismail, Helder Cruz da Silva and Lucas Carvalho Cordeiro}
\affA{Federal University of Amazonas \\
Av. Rodrigo Ot\'avio, 3000, Campus Universit\'ario, Manaus, Brazil, ZIP 69077-000 \\
E-mail: alessandro.b.trindade@gmail.com* \\
E-mail: rdegelo@gmail.com \\
E-mail: esj.galvao@gmail.com \\
E-mail: hussamaismail@gmail.com \\
E-mail: helder@ufam.edu.br \\
E-mail: lucascordeiro@ufam.edu.br \\
* Corresponding author}

\begin{abstract}
Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) is well known by its capability to verify software. However, its use as optimization tool, to solve hardware and software (HW-SW) partitioning problem, is something new. In particular, its integration with the Maximum Satisfiability solver $\nu$Z tool, which provides a portfolio of approaches for solving linear optimization problems over SMT formulas, is unprecedented. We present new alternative approaches to solve the HW-SW partitioning problem. First, we use SMT-based BMC in conjunction with a multi-core support using Open Multi-Processing to create four variants to solve the partitioning problem.
The multi-core SMT-based BMC approaches allow initializing many verification instances based on the number of available processing cores, where each instance checks a different optimum value until the optimization problem is satisfied. Additionally, we integrate the $\nu$Z into the BMC, making it as a specialized solution for optimization in a single-core environment. We implement all five approaches on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC) and compare them to a state-of-the-art optimization tool. Experimental results show that there is no single optimization tool to solve all HW-SW partitioning benchmarks, but based on medium-size benchmarks, ESBMC-$\nu$Z had better performance.

@@

\REF{to this paper should be made as follows: Trindade, A.B., Degelo, R.F., Santos Junior, E.G., Ismail, H.I., Silva, H.C., and Cordeiro, L.C. (2016) `Multi-Core Model Checking and Maximum Satisfiability Applied to Hardware-Software Partitioning', {\it International Journal of Embedded Systems}, Vol. x, No. x, pp.xxx\textendash xxx.}

%Trindade, A., Ismail, H., Galv\~ao, E., Degelo, R., Silva, H., and Cordeiro, L.%
\begin{bio}
Alessandro Bezerra Trindade received the B.Sc. and M.Sc. degrees in electrical engineering from the Federal University of Amazonas (UFAM), in 1995 and 2015, respectively. Currently, he is pursuing his PhD in informatics at UFAM. \vs{9}

\noindent Renato de Faria Degelo, received his software analysis and development technologist degree in 2011 from Uninorte and is pursuing his M.Sc. at UFAM. Currently, he is a software engineer at Samsung Institute (SIDIA).\vs{8}

\noindent Edilson Galvão dos Santos Junior, received his computer engineering degree in 2015 from Foundation Center for Analysis, Research and Technological Innovation (FUCAPI) and is pursuing his M.Sc. at UFAM. Currently, he is a software engineer at SIDIA.\vs{8}

\noindent Hussama Ibrahim Ismail holds a B.Sc. degree in computer engineering from FUCAPI in 2013 and a M.Sc. degree in Electrical Engineering from UFAM in 2015. His current research interests are formal methods and embedded systems.\vs{8}

\noindent Helder Cruz da Silva received the B.Sc. degree and the M.Sc. from the Federal University of Uberlândia (UFU), in 1998 and 2001, respectively. He received the Ph.D. degree from University of Minho in 2013. Currently, he is an assistant professor at UFAM.\vs{8}

\noindent Lucas Carvalho Cordeiro received the Ph.D. degree from the University of Southampton in 2011. From 2009 to 2016, he was an adjunct professor at UFAM (he is currently in unpaid leave). Since 2016, he is a researcher at the University of Oxford.\vs{8}

%\textcolor{red}{\noindent No no no.\vs{8}}

%\textcolor{red}{\noindent No no no.}
\end{bio}

@@

%-----------------related work came to here and was rewriten
Since the first decade of 2000s, two main paths have been tracked to solve the HW-SW partitioning problem: to find the exact solution of a optimization problem, as presented by~\cite{Mann2007}; and to use heuristics to speed up performance time, as presented by~\cite{Arato2003} and~\cite{Arato2005}. It is worth mentioning that using heuristics the final solution is not necessarily an optimal global solution.

There is also the effort to create a hardware-software (HW/SW) co-simulation environment in order to help designers obtain an appropriate HW/SW partitioning that satisfy specified tradeoffs, as described by~\cite{Jan2005} and by~\cite{Hau2009}.

In terms of SMT-based verification, most related studies are restricted to present the model, its modification to specific programming languages ({\it e.g.}, C/C++ and Java), and the application to multi-thread algorithms or to embedded systems to check for program correctness. In~\cite{Ramalho2013} it presents a bounded model checker for C++ programs, which is an evolution of dealing with C programs, and~\cite{Cordeiro2012} use ESBMC for embedded ANSI-C software. In~\cite{Trindade2015}, and~\cite{Trindade2016} it was proven that it is possible to use ESBMC to solve HW-SW partitioning in a single- and multi-core way, but the former has performance issues that were improved by the latter, which used only a sequential search to perform multi-core model checking. There are related studies focused on decreasing the verification time of model checkers by applying Swarm Verification, as mentioned in~\cite{Holzmann2011}, and modifications of internal search engines to add support for parallelism, as presented by~\cite{Holzmann2012}, but there is still the need for initiatives related to parallel SMT solving, as described by~\cite{Wintersteiger2009}.

Recently, the SMT solver Z3 has been extended to pose and solve optimization problems modulo theories, as presented by~\cite{Bjorner2015}. In particular, $\nu$Z tool offers substantial performance improvement in optimization problems, as described by~\cite{Bjorner2014}. Additionally, as an application example, \cite{Pavlinovic2015} propose an approach which considers all possible compiler error sources for statically typed functional programming languages and reports the most useful one subject to some usefulness criterion. The authors formulate this alternative single-core approach as an optimization problem related to SMT and use $\nu$Z to compute an optimal error source in a given ill-typed program.

@@

Here, we apply SMT-based verification methods to the HW-SW partition problem in three different ways using a multi-core ESBMC approach with OpenMP: ESBMC-SS using a sequential-search (SS), ESBMC-PS using a parallel-search (PS), and ESBMC-PB using a binary-search (BS). Our experimental results are compared to ILP (integer linear programming) and GA (generic algorithms) in a multi-core version, and also to $\nu$Z, which supports only a single-core approach, as described by~\cite{Bjorner2015}. The ILP and GA algorithms are implemented with the optimization toolbox of Matlab, as described in~\cite{TheMathWorks2013}, while $\nu$Z is a built-in tool to the SMT solver Z3. All multi-core ESBMC approaches, together with $\nu$Z, are implemented with the ESBMC tool. To the best of our knowledge, this is the first work to use a multi-core SMT-based verification and a MaxSMT solver to check for HW-SW partitioning problems in embedded systems.

@@

Our experiments are based on a set of publicly available benchmarks. All tools, benchmarks, and results of our evaluation are available on a supplementary web page \texttt{http://esbmc.org} and \texttt{http://esbmc.org/benchmarks}.

@@

Optimization is the act of obtaining the best result ({\it i.e.}, the optimal solution) under given circumstances as defined in~\cite{Rao2009}. There is no single method available for efficiently solving all optimization problems, as described by~\cite{Rao2009}. The most well-known technique is linear programming, which is a method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq.~\eqref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors from the objective function, while $Aeq$ and $beq$ are matrices that describe the linear equality constraints

@@

\vspace{2 mm}
Here, we refer to the relevant work of ~\cite{Mann2007} in optimization, who modified a branch-and-bound algorithm to speed up the execution time. ~\cite{Sapienza2013} used multiple criteria decision analysis to solve the HW-SW partitioning problem.

In some cases, the time to find a solution using ILP is impractical. Even with the use of powerful computers, a problem can take hours before an optimal solution is reached. If the optimization problem is complex, some heuristics can be used to solve the same problem faster, as described by~\cite{Rao2009}, {\it e.g.}, those used in the GA. The only drawback is that the found solution may not be the global minimum or maximum. As relevant studies, ~\cite{Jiang2012} used a GA mixed with simulated annealing and~\cite{Huong2012} designed a modified Pareto optimization using GA. In the other hand, there is some research done to reduce the complexity of computation in order to improve the performance of some techniques, as described by~\cite{Boucheneb2006}.

Alternatively, tools such as ESBMC and $\nu$Z can be used to solve optimization problems so that the global minimum or maximum solution is found, as described by~\cite{Trindade2015} and~\cite{Bjorner2015}. The following sections describe the main features of ESBMC and $\nu$Z tools.

@@

Among recent model checking techniques, there is one that combines model checking with satisﬁability solving. This technique, known as bounded model checking (BMC), does a very fast exploration of the state space, and for specific types of problems, it offers large performance improvements over previous approaches, as presented by~\cite{Biere2009}. In particular, BMC based on Boolean Satisfiability (SAT) has been introduced as a complementary technique to binary decision diagrams for alleviating the state-space explosion problem, as described by~\cite{Clarke2001}.

The basic idea of BMC is to check the negation of a given property at a given depth: given a transition system $M$, a property $\phi$, and a bound $k$, BMC unrolls the system $k$ times and translates it into a verification condition (VC) $\psi$  such that $\psi$ is satisfiable if and only if $\phi$ has a counterexample of depth $k$ or less, as defined in~\cite{Biere2009}. To cope with increasing software complexity, SMT solvers can be used as back-ends for solving the generated VCs, as presented by~\cite{Ganai2006},~\cite{Armando2009},~\cite{Cordeiro2012}.

@@

The main factor for ESBMC to use only a single-core relies on its back-end ({\it i.e.}, SMT Solver). Currently, the SMT solvers supported by ESBMC are: Z3, as presented by~\cite{DeMoura2008}; Boolector, as mentioned in~\cite{Brummayer2009}; MathSAT, as presented by~\cite{Barrett2011}; CVC4, as presented by~\cite{Bozzano2005}; and Yices, as presented by~\cite{Dutertre2014}. Most of them do provide neither multi-threaded support nor a parallel version to solve the generated SMT equations.

@@

In particular, OpenMP uses the \textit{fork-join} model of parallel execution, as mentioned in~\cite{OpenMP1998}. The main thread executes the sequential parts of the program; if a parallel region is encountered, then it forks a team of worker threads. After the parallel region finishes ({\it i.e.}, the API waits until all threads terminate), then the main procedure returns to the single-threaded execution mode, as presented by~\cite{Wu2014}.

@@

In this study, the SMT solver Z3 is used to check for the satisfiability of formulas generated from the HW-SW partitioning problem, as presented by~\cite{Bjorner2014}. In particular, we exploit the use of MaxSMT solver $\nu$Z, which is implemented on top of the SMT solver Z3, in order to solve optimization problems; the $\nu$Z main function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.~$\nu$Z also includes an incremental version of the Maximum Resiliency (MaxRes), as mentioned in~\cite{Federica2008}, in order to achieve Maximum Satisfiability (MaxSAT), as defined in~\cite{NarodytskaN} and a Simplex to solve numbers without defined patterns.

In $\nu$Z, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives, as presented by~\cite{Bjorner2015}. In summary, $\nu$Z provides three main functions that extend Z3 for solving optimization problems, which are: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

\begin{itemize}
\item{\textbf{maximize(T)}
this function instructs the solver that a given variable $T$ should be maximized, which includes real, integer, or bit-vector variables.}
\item{\textbf{minimize(T)}
this function instructs the solver that a given variable $T$ should be minimized; the accepted types are the same as maximize function.}
\item{\textbf{Assert-Soft F : weight n}
the function \textit{assert-soft} adds a restriction to $F$, which can also add a weight $n$; the default value is $1$.}
\end{itemize}

@@

\vspace{2 mm}
Fig.~\ref{vZ-Architecture} shows the $\nu$Z architecture. Initially, the SMT formula with objectives is converted to $0-1$ constraints, which leads to a Pseudo-Boolean Optimization (PBO), as mentioned in~\cite{Barth1995} and~\cite{Vasco2005}. If there are many objective functions, $\nu$Z invokes OptSAT for arithmetic or MaxSAT for soft constraints. For constraints using real values, $\nu$Z combines linear arithmetic objectives and uses only one instance of OptSMT. When ``soft constrains'' is used in the mode ``lexicographic'', $\nu$Z invokes MaxSAT using multiple calls for its engine.

@@

The informal model can be described by five characteristics. First, there is only one software context, {\it i.e.}, there is just one general-purpose processor, and there is only one hardware context. The system's components must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the execution time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, or energy consumption; the decision is typically made by the engineer or from the project requirement. Fourth, based on the assumption that hardware is significantly faster than software, the execution time of the components in hardware is considered to be zero. Finally, if two components are mapped to the same context, then there is no communication overhead between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this study. Hardware components do not need scheduling, because the execution time is assumed to be zero. Since there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, with a simplified HW-SW model, where the focus is on bipartitioning, as presented by~\cite{Teich2012}.

@@

A directed simple graph $ G = (V,E) $, called the task graph of the system, is given. Where the vertices $V = \{V_1,V_2,\dotso,V_n\}$ represent the nodes that are the components of the system that will be partitioned. The edges $E$ represent communication between components. Additionally, each node  $V_i$ has a cost $h(V_i)$ (or $h_i$) of hardware if implemented in hardware and a cost $s(V_i)$ (or $ s_i $) of software if implemented in software. Finally, $c(V_i,V_j)$ represents the communication cost between $V_i$ and $V_j$ if they are implemented in different contexts (hardware or software).

@@

\vspace{2 mm}
In particular, different optimization (and decision) problems can be defined for partitioning HW-SW, as described by~\cite{Arato2003}. In this paper, however, the focus is on systems with hard real-time constraints: $ S_0 $ is given (initial cost of software), {\it i.e.}, the objective is to find a $P$ HW-SW partitioning so that $ S_p \leq S_0 $ and $ H_p $ is minimal, which is thus related. Therefore, the objective function is:

@@

Consequently, the constraints can be reformulated based on Equations~\eqref{linear-programming-problem} and~\eqref{software-communication-costs} as:

@@

\vspace{2 mm}
\noindent where $h$, $s$ and $c$ are the vectors representing the cost functions, $E$ is the transposed incidence matrix of $G$ (indicating which edges cross the boundary between the contexts of hardware and software), $n$ is the number of nodes, and $x$ represents the decision variable (a binary vector indicating the partition: $1$ if the node is realized in hardware and $0$ if the node is realized in software). Concerning the complexity of this problem, \cite{Arato2003} demonstrate that it is NP-Hard, as defined in~\cite{Cormem}.

@@

\noindent As computer hardware architecture moves from single- to multi-core, and more recently to heterogeneous computing as described by~\cite{Mittal2015}, parallel programming environments should be exploited to take advantage of the ability to run several threads on different processing cores. This section describes the verification algorithm using sequential ESBMC, followed by three multi-core model checking algorithms and the integration of the MaxSMT solver $\nu$Z into ESBMC, in order to speed up the HW-SW partitioning verification. HW-SW partitioning using ILP-based and Genetic Algorithms are also explained.

@@

Note that, as defined by the formal model, there is an index for each decision variable, indicated by the letter \textit{``i''}, which ranges from $0$ to the number of nodes of the (particular) problem to be solved.

@@

In the ESBMC algorithm, there is no need for adding slack variables in Eq.~\eqref{hw-sw-partitioning}, which reduces the number of variables to be solved if compared to ILP and GA.
